\documentclass[11pt]{article}

% ---------------------------------------------------------------------------
% CS2810 Embodied AI (Spring 2026)
% Combined project report: Goalkeeper + Shooter
% Build: pdflatex main.tex  (run twice for references/links)
% ---------------------------------------------------------------------------

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[strings]{underscore} % allow bare _ in text and \texttt
\usepackage{newtxtext,newtxmath} % professional Times-like body + math
\usepackage[margin=1in]{geometry}
\usepackage{microtype}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{array}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{caption}
\usepackage{enumitem}
\usepackage{listings}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}

\captionsetup{font=small,labelfont=bf}
\setlist{nosep,leftmargin=1.4em}

\lstset{
  basicstyle=\ttfamily\small,
  breaklines=true,
  columns=fullflexible,
  keepspaces=true,
  frame=single,
  framesep=4pt,
  xleftmargin=2pt,
  aboveskip=6pt,belowskip=6pt,
}

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\TODO}[1]{\textcolor{red}{\textbf{[TODO: #1]}}}

\title{\textbf{Learning Shooting and Goalkeeping Skills\\
for the Unitree G1 Humanoid}}
\author{% TODO: add team member names here before submission
CS2810 Embodied AI, Spring 2026}
\date{}

\begin{document}
\maketitle

\begin{abstract}
\noindent
We develop reinforcement-learning soccer skills for the 29-DoF Unitree G1
humanoid in the mjlab MuJoCo simulator, covering the project's two tasks:
\emph{shooting} (approach and kick a stationary ball into the goal) and
\emph{goalkeeping} (intercept a ball launched on a parabolic trajectory). For
the shooter we build a six-stage progressive perception--action pipeline on a
single shared LSTM policy, reaching a \textbf{99.96\%} goal rate over
$3\times10^4$ trials (Phase-1 score $\approx 29.9/30$) with $12$--$16$\,m/s ball
speeds, and deploy it behind a multi-strategy API server whose
goalkeeper-aware policy reads the opponent and shoots to the far side. For the
goalkeeper we treat diving saves -- which are unreachable by reward-shaped
exploration and suppressed by any upright prior -- as a sequence of targeted
sub-problems: distil a diving base from a reference teacher, \emph{prove} with a
per-scenario CEM repair oracle that conceded high balls are physically saveable,
and distil the repaired saves back into a single deployable MLP. This lifts the
goalkeeper block rate from a $66.9\%$ distilled base to \textbf{81.7\%}. We
report both tasks honestly, including a clean negative result on on-policy
ballistic-residual fine-tuning, and detail the API observation computation that
each policy performs server-side from raw MuJoCo state.
\end{abstract}

\setcounter{tocdepth}{1}
{\small\tableofcontents}
\medskip

% ===========================================================================
\section{Introduction}
% ===========================================================================

The project develops two humanoid soccer skills on the Unitree G1 (29 DoF) in
the mjlab MuJoCo-based physics simulation~\cite{mujoco}, each controlled at
$50$\,Hz:

\begin{itemize}
\item \textbf{Shooting}: the robot approaches a stationary ball and kicks it
through the goal frame. This is a perception-guided motor skill that needs a
motion-tracking prior and ball-trajectory control.
\item \textbf{Goalkeeping}: the robot intercepts an incoming ball launched on a
parabolic trajectory toward a $3.0\,\text{m}\times1.8\,\text{m}$ goal. This is a
reactive interception task requiring rapid end-effector positioning and ball
trajectory estimation.
\end{itemize}

The grading has two phases. \textbf{Phase~1} evaluates each single-agent policy
over $50$ trials with the official \code{eval_naive_shooter.py} /
\code{eval_naive_goalkeeper.py} scripts, scored by the linear band
$\text{score}=\max\!\big(0,\,(\text{success}-0.8)/0.2\big)\times 30$ (zero below
$80\%$, full at $100\%$). \textbf{Phase~2} is a centrally run cross-evaluation
tournament in which each team serves its policy behind a standard
\code{/act}\,/\,\code{/reset} HTTP API and \code{compete.py} queries both teams
at every simulation step. Recent humanoid-soccer work informs the environment
and training design~\cite{kong,ren}.

This report is organized by task. Part~\ref{part:keeper} (Sections
\ref{sec:gk-task}--\ref{sec:gk-api}) covers the goalkeeper; Part~\ref{part:shooter}
(Sections \ref{sec:sh-overview}--\ref{sec:sh-engineering}) covers the shooter.
Each part reports its Phase-1 results with analysis and its Phase-2 design,
including how observations are computed from raw state.

% ===========================================================================
\part*{Part A\quad Goalkeeper}\label{part:keeper}
\addcontentsline{toc}{section}{Part A: Goalkeeper}
% ===========================================================================

The deliverable is the \code{keeper-lyk} branch: a repair-oracle plus
distillation pipeline that lifts a distilled MLP goalkeeper to an \textbf{81.7\%}
block rate (checkpoint \code{src/assets/soccer/weight/model_repaired_lyk.pt}).

% ---------------------------------------------------------------------------
\section{Task and the core difficulty}\label{sec:gk-task}
% ---------------------------------------------------------------------------

The goalkeeper must intercept a ball launched on a parabolic trajectory toward a
full $3\,\text{m}\times1.8\,\text{m}$ goal, controlling the G1 at $50$\,Hz. A
save is credited only if the ball never crosses the goal plane within the frame.
Crucially, the diving saves required to defend the corners necessarily terminate
with the robot on the ground.

This single property reshapes the entire problem. Any learning signal that
penalizes falling simultaneously penalizes the dives that produce saves, so the
conventional ``remain upright'' inductive bias is actively counterproductive; the
evaluation reflects this by not treating a fall as a failure. The consequence is
that \textbf{the required interception behavior is not reachable by reward-shaped
exploration} -- under a hand-designed reward, a humanoid does not stumble into a
committed, directional dive through random search. We verified this empirically:
trained from scratch with PPO~\cite{ppo} under a range of shaped rewards, the
policy reliably learns to stand and to satisfy the proxy terms, but its block
rate saturates between $3\%$ and $33\%$, with the high-ball corners frozen near
zero. A weak reward yields passive standing ($\approx 33\%$, barely above the
do-nothing baseline); a strong one induces an uncontrolled lunge that topples
without intercepting.

The central design question is therefore not ``which reward yields a goalkeeper''
but rather \textbf{how to inject a diving prior, and how to subsequently improve
upon it against the true objective}. The remainder of the method follows from
this premise.

A second observation establishes the target. A per-region diagnosis of the
Humanoid Goalkeeper of Ren et al.~\cite{ren} -- whose checkpoint serves as our
teacher -- shows that it blocks only $\approx 72.5\%$ on our benchmark, with
failures concentrated almost entirely in the high balls (low balls
$\approx 95\%$, the upper corners $\approx 40$--$60\%$). This is doubly
informative: competent behavior still leaves the corners largely undefended, and
it implies that \textbf{pure imitation of Ren et al.~\cite{ren} cannot exceed
$\approx 72.5\%$}. To do better, we must \emph{synthesize} high-ball saving
behavior that the teacher does not itself possess.

% ---------------------------------------------------------------------------
\section{Method: a repair-oracle distillation pipeline}\label{sec:gk-method}
% ---------------------------------------------------------------------------

Because the diving prior cannot be acquired from scratch and imitation of
Ren et al.~\cite{ren} is upper-bounded at $\approx 72.5\%$, no single training
procedure is sufficient. We adopt a \textbf{staged pipeline} in which each stage
overcomes the failure mode of the preceding one: distil a diving base,
\emph{prove} with an offline optimizer that the conceded balls are physically
saveable, and then fold those repaired saves back into a single deployable
policy by distillation.

\subsection{Stage A --- Distillation: injecting the diving prior}

We behavior-clone the teacher policy of Ren et al.~\cite{ren} into a trainable
network of our own design, since the released checkpoint is inference-only under
the current RL framework~\cite{rudin} and cannot itself be optimized. The
procedure follows DAgger~\cite{dagger}: the student is rolled out, the teacher
labels the visited states, and the aggregated dataset is cloned. This yields a
competent, reactive \emph{diving} base policy -- the $960$-D-input MLP checkpoint
\code{goalkeeper_distilled_v3.pt}. It does not surpass the teacher -- imitation
cannot -- scoring $\approx 74\%$ over $100$ trials and $\approx \mathbf{66.9\%}$
on the harder vectorized benchmark used in the rest of this part. It nevertheless
provides a trainable policy that already executes dives, the precondition for
every subsequent stage.

\subsection{Stage B --- Repair oracle: proving the hard balls are saveable}

To exceed the base policy we must save high balls that it concedes. We construct
a \textbf{per-scenario trajectory optimizer}: for the frozen base policy and a
fixed incoming ball, it searches -- using the cross-entropy method,
CEM~\cite{cem,icem}, a sampling-based optimizer -- an open-loop \emph{residual}
action sequence added to the base policy's actions. The search minimizes a dense
cost combining the true concede outcome with the distance from the nearest
blocking limb to the ball's predicted plane-crossing point. This optimizer
\textbf{demonstrates that the hard balls are physically saveable} at a
per-scenario save rate far above the base policy, establishing that control
\emph{authority} rather than kinematic reach was the limiting factor. The
optimizer is single-agent simulation optimization, not a learned policy -- it is
used as a \emph{data generator}, not as the deliverable.

\subsection{Stage C --- Distilling the repaired saves into a deployable MLP}

The repaired residuals are per-scenario and open-loop; they cannot be deployed
directly (Section~\ref{sec:gk-results}). We therefore \textbf{collect the
repaired \code{(observation, action)} pairs from trajectories that the oracle
turns into successful saves, and distil them back into a native MLP} with the
same architecture as the base. The distilled checkpoint,
\code{model_repaired_lyk.pt}, reaches \textbf{81.7\%} -- a $+14.8$-point gain
over the $66.9\%$ base. Two collection choices proved essential, each of which
silently degraded the result when violated:

\begin{itemize}
\item \emph{Demonstrations must be evaluation-matched.} Forcing a scenario after
the environment reset leaves stale ball-history frames in the observation,
training the policy on a distribution it never encounters at evaluation. The
scenario must be imposed \emph{through} the reset so that the observation history
is consistent with the deployed keeper observation/history layout.
\item \emph{Collect a stable save window, not the post-save fall.} A naive
collection learns the high-ball save together with the subsequent
topple-and-flail, so the deployed policy blocks but then loses balance after
contact. The oracle residual is faded back to the base policy after the save
window, and distillation keeps \textbf{only the frames around the keeper-plane
crossing instant} (a short pre/post window), so the student learns the
interception rather than the fall.
\end{itemize}

A \textbf{region-weighted, multi-GPU collection} path supports the same pipeline:
one repair worker per GPU, with region weights that oversample low/mid balls
(regions $4,5$ low, $0,1$ mid, $2,3$ upper) to keep the distilled policy from
overfitting to the dramatic high-ball jumps at the expense of routine balls.

% ---------------------------------------------------------------------------
\section{Phase 1 result and analysis}\label{sec:gk-results}
% ---------------------------------------------------------------------------

\paragraph{Evaluation protocol.} All numbers use the official grader script,
\code{scripts/eval_naive_goalkeeper.py} on the \code{Eval-Goalkeeper} task -- a
6-region parabolic ball, success defined exactly as the rubric requires (the
ball does not cross the goal plane inside the frame), with \code{fell_over}
termination disabled so a diving save that ends on the ground still counts. The
deliverable checkpoint is \code{model_repaired_lyk.pt}, reproduced with:

\begin{lstlisting}
MUJOCO_GL=egl python scripts/eval_naive_goalkeeper.py \
  --headless --num-trials 50 \
  --checkpoint src/assets/soccer/weight/model_repaired_lyk.pt
\end{lstlisting}

\paragraph{Block rate: 81.7\%}, up from the $\mathbf{66.9\%}$ distilled base
($+14.8$ points) and above the $\approx 72.5\%$ reference teacher of
Ren et al.~\cite{ren} on the identical benchmark, obtained without an unrealistic
robot model and without deploying the released checkpoint as-is. The $81.7\%$
figure is the stable estimate over a large trial count; it is reported as the
trustworthy number because at the rubric's \textbf{50-trial} protocol the
binomial standard deviation is $\approx\pm 5.5$ percentage points, so a single
50-trial grader run scatters around this mean. Executing the reproduce command
above on the GPU server gives an \textbf{official 50-trial result of
$42/50 = 84.0\%$} (the loaded actor input is confirmed at $960$-D). Under the
rubric's linear band -- zero at $80\%$, full marks at $100\%$ -- this maps to a
score of $(0.84-0.80)/0.20\times 30 = 6.0/30$, versus $\approx 2.5/30$ for the
$81.7\%$ central estimate. The single 50-trial run (the official script exposes
no seed flag) and the $81.7\%$ large-sample figure agree to within the
$\approx\pm 5.5$ percentage-point binomial deviation of a 50-trial measurement;
we report $81.7\%$ as the trustworthy central estimate and $84.0\%$ as the
observed grader-protocol result. Either way the margin above the $80\%$ scoring
floor is real but modest, which is the honest characterization of this
deliverable.

\paragraph{Residual failure structure.} Failures remain concentrated in the high
balls. The policy still \emph{overuses} high-ball jump saves and can lose balance
after contact on the hardest upper-corner trajectories; the low and mid regions
are comfortably defended. The per-region picture mirrors the teacher's: low balls
are near-solved, and the headroom is in the upper corners, where the dive must
arrive within a few centimeters of the crossing point at exactly the crossing
instant.

\paragraph{A negative result on the oracle, for completeness.} The repair oracle
of Stage~B reaches a much higher save rate \emph{per ball}, but this is a
per-scenario optimum rather than deployable headroom: its residual is
co-optimized with the base policy's exact trajectory for one specific ball, so
replaying or retrieving those open-loop dives at inference scores \emph{below}
the base policy, because on any other ball the fixed dive conflicts with the
reactive base. The deployable gain therefore had to come from \textbf{distilling}
the repaired saves into a reactive policy (Stage~C), not from executing oracle
trajectories directly -- and a gap remains between the per-scenario oracle and
the single reactive network that absorbs it, which is the subject of
Section~\ref{sec:gk-secondary}.

% ---------------------------------------------------------------------------
\section{Secondary experiments and negative results}\label{sec:gk-secondary}
% ---------------------------------------------------------------------------

Two further directions were explored on this branch. Both are reported honestly:
one is a clean negative, the other a partially-validated path to close the
oracle-to-policy gap.

\paragraph{Ballistic-residual PPO (negative).} A natural hypothesis is that the
base policy misses hard balls only because it must \emph{infer} the ball's future
crossing point from history, so giving it explicit ballistic features should
help. We froze the distilled $960$-D MLP and trained only a small, bounded,
zero-initialized residual head (\code{GoalkeeperBallisticResidual}) with
PPO~\cite{ppo}, feeding it explicit timing features: time and $(y,z)$ at the
keeper plane $x=0$, time and $(y,z)$ at the goal plane $x=-0.5$, the incoming
$v_x$, and the ball speed. Because the residual is zero-initialized, training
starts exactly at the base policy -- but early 2048-env runs \emph{stayed} near
the frozen base and produced no clear gain. The conservative residual is safe but
does not, on its own, synthesize the hard-ball saves; the productive signal lives
in the repair oracle, not in the on-policy residual.

\paragraph{Residual-target distillation (gap-closer).} Distilling the repaired
saves into a \emph{full-MLP replacement} (Stage~C) reaches $81.7\%$, well short
of the oracle's per-ball ceiling. To narrow that gap without destabilizing the
base, the same repair data can instead be distilled into the
\textbf{frozen-base ballistic-residual} form -- keeping the strong base intact and
learning only the bounded corrective head from the oracle demonstrations rather
than from on-policy reward. This couples the one component that demonstrably
synthesizes hard-ball behavior (the oracle) with the one that deploys safely (the
frozen base plus bounded residual), and is the recommended next checkpoint for
pushing past $81.7\%$.

The conclusion is that, for this task, the gains reside in the
\textbf{repair-oracle data generation plus distillation} pipeline rather than in
on-policy fine-tuning of the reactive policy, and we retain the repaired
distilled MLP as the deliverable.

% ---------------------------------------------------------------------------
\section{Phase 2: deployment behind the competition API}\label{sec:gk-api}
% ---------------------------------------------------------------------------

For the cross-evaluation tournament the policy is served behind the standardized
\code{/act} and \code{/reset} HTTP API (\code{scripts/api_server.py}). The
grader's \code{compete.py} sends the \textbf{same raw MuJoCo state} to every
team -- root pose/velocity, joint positions/velocities and last action for both
robots, plus the ball position and velocity -- and the server turns that raw
state into exactly the observation tensor the policy was trained on. The match
between this server-side observation and the training observation is what the
entire repair pipeline depends on, so we specify it concretely.

\paragraph{Per-frame observation (96-D), computed server-side from raw state.}
On each \code{/act} call \code{compute_goalkeeper_obs} builds one $96$-D frame by
concatenating six terms, each transformed into the goalkeeper's own pelvis frame
with the exact scalings used in training (Table~\ref{tab:gk-obs}). Only the ball
position and the keeper's own proprioception are used; no privileged or opponent
state enters the observation. The key non-obvious detail is that the reference
frame and the per-term scalings ($0.25$, $0.05$, the GK default joint offset) are
goalkeeper-specific and \textbf{must} match the training config -- a wrong scale
or a world-frame ball vector silently destroys the policy at deployment even
though the network loads fine.

\begin{table}[h]
\centering\small
\caption{Goalkeeper per-frame observation (96-D) built from raw state.}
\label{tab:gk-obs}
\begin{tabular}{@{}llp{8.3cm}@{}}
\toprule
Term & Dim & Computation from raw state \\
\midrule
\code{ball_pos_local}    & 3  & ball position relative to pelvis: \code{quat_apply(quat_inv(root_quat), ball_pos - root_pos)} \\
\code{base_ang_vel}      & 3  & root angular velocity rotated into body frame, scaled $\times 0.25$ (the GK PD-gain ratio) \\
\code{projected_gravity} & 3  & gravity $(0,0,-1)$ rotated into body frame via \code{quat_apply(quat_inv(root_quat), g)} \\
\code{joint_pos_rel}     & 29 & \code{joint_pos - GK_default_joint_pos} (scale $\times 1.0$) \\
\code{joint_vel_scaled}  & 29 & \code{joint_vel} $\times 0.05$ \\
\code{last_action}       & 29 & the previous action returned by the server \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Term-major history stacking ($\to$ 960-D).} The actor consumes a
$10$-frame history, but \textbf{not} as a naive frame-major concatenation.
\code{stack_goalkeeper_history} reproduces mjlab's \emph{term-major} layout: it
concatenates all $10$ frames of \code{ball_pos_local} first, then all $10$ frames
of \code{base_ang_vel}, and so on per term, yielding the $96\times 10 = 960$-D
vector the MLP expects. The server keeps a $10$-slot history buffer; on the first
frame after a reset it pre-fills all $10$ slots with the initial frame so the
very first action is well-defined.

\paragraph{Endpoints.} \code{/act} computes the frame, updates the history,
term-major-stacks it, and runs the policy under \code{inference_mode}; for the
residual / gate-based checkpoint variants it also forwards the raw ball position
and velocity to the policy via \code{set_raw_ball_state} before the forward pass,
so the ballistic features are computed from ground-truth ball state rather than
re-estimated. \code{/reset} clears the policy's internal state and the history
buffer. The repaired checkpoint additionally stores \code{ballistic_residual} /
repair metadata, so \code{_load_policy} can \textbf{rebuild the exact inference
graph} -- frozen base path plus residual head when present -- from the checkpoint
alone, which is why the same server code serves the plain distilled MLP, the
residual variant, and the recurrent variants without modification.

% ===========================================================================
\part*{Part B\quad Shooter}\label{part:shooter}
\addcontentsline{toc}{section}{Part B: Shooter}
% ===========================================================================

% ---------------------------------------------------------------------------
\section{Overview}\label{sec:sh-overview}
% ---------------------------------------------------------------------------

\textbf{Task.} The G1 humanoid (29 DoF) performs a penalty-spot shot in MuJoCo:
approach a stationary ball and kick it into the goal. This is a
perception-guided motor skill requiring a motion-tracking prior and
ball-trajectory control. Table~\ref{tab:sh-deliverables} summarizes the
deliverables.

\begin{table}[h]
\centering\small
\caption{Shooter deliverables.}
\label{tab:sh-deliverables}
\begin{tabular}{@{}lp{9.5cm}@{}}
\toprule
Artifact & Description \\
\midrule
Six-stage progressive training pipeline & Stage 1 $\to$ Stage 6, 20K iterations, $\approx$ 39B steps \\
Stage-6 model & LSTM 128-64-32 $+$ 2$\times$128, actor 160-D / critic 298-D \\
Multi-strategy API server & \code{uniform} / \code{best-motion} / \code{gk-aware} shooting strategies \\
Evaluation tooling & per-motion analysis, kick-timing analysis, parallel batched eval \\
Performance matrix & per-motion accuracy/speed data over 250K trials \\
\bottomrule
\end{tabular}
\end{table}

% ---------------------------------------------------------------------------
\section{Training pipeline: six-stage progressive training}\label{sec:sh-pipeline}
% ---------------------------------------------------------------------------

Following the progressive perception--action framework of PAiD~\cite{kong}, we
design six training stages. All stages use the same LSTM architecture; capability
is grown purely through the environment configuration (rewards, terminations,
coordinate system), summarized in Table~\ref{tab:sh-stages}.

\begin{table}[h]
\centering\small
\caption{The six training stages.}
\label{tab:sh-stages}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}llll@{}}
\toprule
Stage & Goal & Coordinates & Key new rewards \\
\midrule
\textbf{1} & Motion tracking (imitate kick motions) & Training & six \code{track_body} terms, pure imitation \\
\textbf{2} & Perception-guided kicking & Training & \code{contact} $+$ \code{sideways_kick} $+$ \code{ball_vel_align} \\
\textbf{3} & Goal-plane accuracy $+$ speed curriculum & Training & \code{goal_accuracy} $+$ \code{goal_miss} \\
\textbf{4} & High-speed accuracy (target 10\,m/s) & Training & \code{is_terminated}$=-500$, \code{ball_speed}$=20$ \\
\textbf{5} & Competition-frame alignment $+$ full-goal accuracy & \textbf{Competition} & \code{goal_accuracy}$=30$, \code{ball_oob}$=-30$ \\
\textbf{6} & High-speed kicking in competition frame & Competition & \textbf{\code{ball_speed}$=40$}, \code{z_speed}$=-5$ \\
\bottomrule
\end{tabular}}
\end{table}

\subsection{Coordinate systems}
\begin{itemize}
\item \textbf{Training frame (Stages 1--4):} robot near the origin, goal along
$-y$. \code{motion_origin_offset} $+$ \code{motion_yaw_offset} map motion-local
coordinates into world coordinates.
\item \textbf{Competition frame (Stages 5--6):} fully aligned with
\code{compete.py} -- robot at $(4,0,0.8)$ facing $-x$, ball at $(3,0,0.1)$, goal
plane $x=-0.5$, width $3.0$\,m, height $1.8$\,m.
\end{itemize}

\subsection{Training environment parameters}
\begin{table}[h]
\centering\small
\caption{Training environment parameters.}
\begin{tabular}{@{}ll@{}}
\toprule
Parameter & Value \\
\midrule
Parallel environments & 4096 \\
Steps per environment & 24 \\
Control frequency & 50\,Hz (decimation $=4$, timestep $=0.005$\,s) \\
Episode length & 10.0\,s \\
Action space & 29-D joint position targets \\
Action scaling & \code{G1_ACTION_SCALE} (per-joint PD gains) \\
PPO learning rate & $1\times10^{-3}$ (adaptive schedule) \\
Discount $\gamma$ & 0.99 \\
GAE $\lambda$ & 0.95 \\
KL target & 0.01 \\
Entropy coef. & 0.005 \\
GPU & 2$\times$ NVIDIA GeForce RTX 4090 D (24\,GB) \\
Training time (Stage 6) & $\approx 9$\,h \\
Total steps (Stage 6) & $\approx 3.93\times10^{9}$ \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Per-stage detail}

\paragraph{Stage 1 --- Motion tracking.} Learn to imitate diverse kick-motion
references with no perception input, forming a stable motion prior.
\emph{Command sampling:} adaptive (biased toward high-failure motion segments;
kernel size 3, $\lambda=0.1$, uniform ratio $0.1$). \emph{Domain randomization:}
\code{push_robot} (random base-velocity impulse every 1--3\,s), observation noise
($\pm0.01$ to $\pm0.5$ per term), observation corruption. \emph{Terminations}
(5): \code{time_out}; \code{fell_over} (base tilt $>70^{\circ}$); \code{anchor_pos_z}
($z$ error $>0.25$\,m); \code{anchor_ori} (orientation error $>0.8$);
\code{ee_body_pos} (end-effector position error $>0.25$\,m). The last three are
motion-reference guards.

\paragraph{Stage 2 --- Perception-guided kicking.} Introduce ball perception and
soccer kick rewards; relax motion constraints to allow ball chasing. Command
sampling switches from adaptive to uniform (each episode starts at frame 0). New
rewards: \code{proximity} (1.0), \code{contact} (50.0, foot--ball contact with
horizontal force), \code{sideways_kick} (50.0, ball moves toward goal),
\code{ball_vel_align} (30.0), \code{ball_speed} (10.0), \code{foot_distance}
($0.2$), \code{is_terminated} ($-200$), and cheat suppressors
\code{both_feet_ball} ($-5$), \code{nonfoot_ball} ($-3$), \code{foot_stomp}
($-20$), \code{foot_lift} ($-2$). Motion-tracking weight \code{anchor_pos} drops
to $0.0$ (free chasing); \code{track_foot_pos}$=1.0$ added. \code{ee_body_pos}
threshold relaxed $0.25\to0.35$\,m.

\paragraph{Stage 3 --- Goal-plane accuracy $+$ speed curriculum.} Add an explicit
goal-plane crossing reward and adaptive target sampling. \code{ball_speed}$=15$
(std widened to $3.0$); \textbf{new} \code{goal_accuracy}$=10$ (Gaussian,
std$=0.3$); \textbf{new} \code{goal_miss}$=-5$ (ball never reaches the goal
plane); the six motion-tracking terms halve $1.0\to0.5$. Terminations are sharply
pruned to just \code{time_out} $+$ \code{fell_over} (all motion-reference guards
removed). Adaptive target: 11 goal-plane bins, $\alpha=0.3$ (focus on the
hardest-to-hit regions).

\paragraph{Stage 4 --- High-speed accuracy.} Push ball speed toward
$\approx 10$\,m/s. \code{ball_speed}$=20$ (std $5.0$); \code{is_terminated}
$-200\to-500$ (a fall is more costly at speed); \code{undesired_contacts}
$-0.1\to-1.0$ (prevent body dragging); \code{goal_miss} $-5\to-20$.

\paragraph{Stage 5 --- Competition-frame alignment.} \textbf{Key innovation:}
switch the training frame entirely to \code{compete.py}'s world coordinates,
eliminating the train$\to$compete domain gap -- the model now sees exactly the
match scene during training. \code{goal_accuracy} $10\to30$;
\code{goal_miss}$\to$\code{goal_miss_scaled} ($-50$, scaled by lateral miss
distance); \code{ball_vel_align}$=30$ (std tightened to $0.12$); \code{z_speed}
re-enabled ($-2$, mild anti-lob); \textbf{new} \code{ball_oob}$=-30$ (ball leaves
the goal bounds); \textbf{new} \code{both_feet_air_time}$=-5$ (both feet airborne
$>0.15$\,s); \code{proximity} $1.0\to3.0$; the six motion-tracking terms drop to
$\approx0.05$ each (total $0.3$). \emph{Domain randomization: fully off} -- zero
perturbation, noise, corruption, or push -- a pure competition environment.

\paragraph{Stage 6 --- High-speed competition.} Push ball speed toward
$12$--$16$\,m/s in the competition frame. \code{ball_speed} $20\to\mathbf{40}$
(std $10.0$); \code{z_speed} $-2\to\mathbf{-5}$ (anti-lob at speed);
\code{ball_vel_align} $30\to20$ (std relaxed to $0.20$); \code{goal_accuracy}
$30\to20$ -- trading accuracy for speed.

% ---------------------------------------------------------------------------
\section{Model architecture and observation space}\label{sec:sh-arch}
% ---------------------------------------------------------------------------

\subsection{Network}
\begin{lstlisting}
RNNModel (Actor & Critic share the architecture):
  MLP encoder: Linear(128->128) -> ELU -> Linear(128->64) -> ELU -> Linear(64->32) -> ELU
  RNN:         LSTM(input_dim, hidden=128, num_layers=2)
  Output head: Linear(32->output_dim)

Actor:  input_dim=160, output_dim=29 (joint position targets)
Critic: input_dim=298, output_dim=1  (value)
\end{lstlisting}
Action distribution: \code{GaussianDistribution} with a scalar (single learnable
log-std), initial std $=1.0$. Observation normalization:
\code{EmpiricalNormalization} on both actor and critic.

\subsection{Actor observation space (160-D)}
\begin{table}[h]
\centering\small
\caption{Actor observation terms (160-D).}
\begin{tabular}{@{}clcp{7.2cm}@{}}
\toprule
\# & Term & Dim & Description \\
\midrule
0 & \code{command} & 58 & reference joint positions (29) $+$ velocities (29) from the motion command \\
1 & \code{projected_gravity} & 3 & gravity vector in the base frame \\
2 & \code{motion_ref_ang_vel} & 3 & reference torso angular velocity (world frame) \\
3 & \code{base_ang_vel} & 3 & measured base angular velocity (IMU gyro) \\
4 & \code{joint_pos} & 29 & current joint position $-$ \code{HOME_KEYFRAME} default \\
5 & \code{joint_vel} & 29 & current joint velocity \\
6 & \code{actions} & 29 & previous-step action output \\
7 & \code{target_point_pos} & 3 & ball position in the robot pelvis frame \\
8 & \code{target_destination_pos} & 3 & shot target point in the pelvis frame \\
\bottomrule
\end{tabular}
\end{table}
All terms are normalized by their per-term std before concatenation.

\subsection{Critic privileged information (+138-D)}
These five extra terms are available only during training (privileged
information), giving the critic ground-truth state for more accurate value
estimation: \code{motion_anchor_pos_b} (3), \code{motion_anchor_ori_b} (6, 6-D
rotation), \code{body_pos} (42, 14 tracked bodies in anchor frame),
\code{body_ori} (84, 14 bodies, 6-D rotation), \code{base_lin_vel} (3, world
frame).

\subsection{Domain randomization schedule}
Domain randomization contracts monotonically from Stage 1 to Stage 6
(Table~\ref{tab:sh-dr}). The fully randomization-free Stage 5--6 design ensures
exact consistency with the \code{compete.py} environment.

\begin{table}[h]
\centering\small
\caption{Domain-randomization schedule.}
\label{tab:sh-dr}
\begin{tabular}{@{}llcccc@{}}
\toprule
Stage & Sampling & Push robot & Obs noise & Obs corruption & Ball randomization \\
\midrule
1   & \textbf{adaptive} (failure-biased) & yes & yes (actor) & yes & --- \\
2--4 & uniform & yes & yes (actor) & yes & curve offset $\pm0.15$\,m \\
5--6 & uniform & \textbf{no} & \textbf{no} & \textbf{no} & \textbf{no} \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Motion files}
Ten \code{soccer-standard-*.npz} motion files are used: six right-leg kicks
(\code{001/004/005/006/009/010_right}) and four left-leg kicks
(\code{002/003/007/008_left}). Each file stores joint positions $(T,29)$, joint
velocities $(T,29)$, and torso angular velocity $(T,3)$; lengths 205--289 frames
(4.1--5.8\,s).

% ---------------------------------------------------------------------------
\section{Reward design evolution}\label{sec:sh-reward}
% ---------------------------------------------------------------------------

There are 28 reward terms at Stage 6, but the core signal is driven by a few
high-weight terms. The design intent is a gradual transition from \emph{pure
imitation} to \emph{pure match performance}
(Table~\ref{tab:sh-reward-evolution}).

\begin{table}[h]
\centering\footnotesize
\caption{Core reward-term evolution across stages.}
\label{tab:sh-reward-evolution}
\setlength{\tabcolsep}{4pt}
\begin{tabular}{@{}lccccccl@{}}
\toprule
Reward term & S1 & S2 & S3 & S4 & S5 & S6 & Function \\
\midrule
\code{track_body_pos}     & 1.0 & 1.0 & 0.5 & 0.5 & \textbf{0.05} & 0.05 & body pose tracking \\
\code{track_body_ori}     & 1.0 & 1.0 & 0.5 & 0.5 & \textbf{0.05} & 0.05 & body orientation tracking \\
\code{track_anchor_ori}   & 1.0 & 1.0 & 0.5 & 0.5 & \textbf{0.05} & 0.05 & anchor orientation tracking \\
\code{track_body_lin_vel} & 1.0 & 1.0 & 0.5 & 0.5 & \textbf{0.05} & 0.05 & body linear-velocity tracking \\
\code{track_body_ang_vel} & 1.0 & 1.0 & 0.5 & 0.5 & \textbf{0.05} & 0.05 & body angular-velocity tracking \\
\code{track_foot_pos}     & --- & 1.0 & 0.5 & 0.5 & \textbf{0.05} & 0.05 & foot position tracking \\
\textbf{tracking total}   & \textbf{6.0} & 5.0 & 3.0 & 3.0 & \textbf{0.3} & \textbf{0.3} & \\
\code{contact}            & --- & \textbf{50} & 50 & 50 & 50 & 50 & foot--ball contact force \\
\code{sideways_kick}      & --- & \textbf{50} & 50 & 50 & 50 & 50 & lateral kick toward goal \\
\code{ball_vel_align}     & --- & 30 & 30 & 30 & 30 & \textbf{20} & ball velocity aligned to goal \\
\code{ball_speed}         & --- & 10 & \textbf{15} & \textbf{20} & 20 & \textbf{40} & ball horizontal speed \\
\code{goal_accuracy}      & --- & --- & \textbf{10} & 10 & \textbf{30} & 20 & goal-plane crossing accuracy \\
\code{goal_miss}          & --- & --- & $-5$ & $-20$ & \textbf{$-50$} & $-50$ & miss penalty \\
\code{z_speed}            & --- & 0 & 0 & 0 & $-2$ & \textbf{$-5$} & vertical-speed penalty \\
\code{ball_oob}           & --- & --- & --- & --- & $-30$ & $-30$ & out-of-bounds penalty \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Regularization (shared across stages).} \code{action_rate} ($-0.1$,
$L_2$ on action delta); \code{joint_limit} ($-10.0$); \code{undesired_contacts}
($-0.1$/$-1.0$, non-foot / non-wrist ground contact).

\paragraph{Design logic.} (1) \emph{Stage 1 (pure imitation):} six tracking
terms total $6.0$, no soccer reward, forming a stable kick prior. (2)
\emph{Stage 2 (add soccer):} sparse rewards (\code{contact}$=50$,
\code{sideways_kick}$=50$) drive foot--ball interaction while negative terms
suppress cheating; tracking eased (\code{anchor_pos}$\to0$). (3)
\emph{Stages 3--4 (accuracy $+$ speed):} dense goal-accuracy reward added,
\code{ball_speed} climbs, \code{goal_miss} hardens $-5\to-20\to-50$, tracking
halved (total $3.0$), terminations pruned $5\to2$. (4) \emph{Stages 5--6
(competition alignment $+$ extreme speed):} tracking near zero ($0.3$),
\code{ball_speed} doubles $20\to40$ to become the main signal, \code{z_speed}
hardens to $-5$, accuracy signals slightly lowered (\code{ball_vel_align}
$30\to20$, \code{goal_accuracy} $30\to20$) to trade accuracy for speed.

% ---------------------------------------------------------------------------
\section{Phase 1 evaluation}\label{sec:sh-phase1}
% ---------------------------------------------------------------------------

\subsection{Setup}
Official grader \code{eval_naive_shooter.py} (single-env, task \code{Eval-Shooter})
and a parallel \code{eval_shooter_parallel.py} (task \code{Eval-Shooter-Stage6},
4096 envs). Model: Stage 6 \code{model_20000.pt}. Seeds: $42, 2810, 202606$.
Trials/seed: $50$ (naive) / $10000$ (parallel). Score formula
$\max(0,(\text{success}-0.8)/0.2)\times30$.

\subsection{Results}
\begin{table}[h]
\centering\small
\caption{Shooter Phase-1 results (parallel eval, 10000 trials/seed).}
\begin{tabular}{@{}lccc@{}}
\toprule
Seed & Trials & Goals & Success rate \\
\midrule
42      & 10000 & 9996 & \textbf{99.96\%} \\
2810    & 10000 & 9995 & \textbf{99.95\%} \\
202606  & 10000 & 9997 & \textbf{99.97\%} \\
\textbf{Mean} & \textbf{10000} & \textbf{9996.0} & \textbf{99.96\%} \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Phase-1 score:} $\max(0,(0.9996-0.8)/0.2)\times30 = 29.9/30$.
Representative metrics (seed 202606): mean kick speed $\mathbf{13.03}$\,m/s, kick
accuracy (cosine similarity) $0.9787\pm0.0191$. The goal rate exceeds $99.9\%$ at
all three seeds (only 3--5 misses per 10000); the $0.1$-point shortfall from full
marks is the linear map magnifying the tiny gap from $100\%$.

\subsection{Evaluation tooling}
To meet the fixed-seed requirement we hardened the eval scripts: (1) four-level
seed control (\code{random} / \code{np.random} / \code{torch.manual_seed} /
\code{env_cfg.seed}); (2) a \code{--seeds} argument running multiple seeds and
auto-aggregating the mean and Phase-1 score; (3) per-trial \code{policy.reset()}
to clear LSTM hidden state; (4) explicit GPU memory management between seeds
(\code{del} $+$ \code{gc.collect()} $+$ \code{torch.cuda.empty_cache()}); (5) a
parallel evaluator giving $\approx25\times$ speedup over 4096 envs with identical
goal-detection logic.

% ---------------------------------------------------------------------------
\section{Phase 2: competition strategy and API server}\label{sec:sh-phase2}
% ---------------------------------------------------------------------------

\subsection{API server}
A FastAPI server loads the Stage-6 LSTM policy (task \code{Eval-Shooter-Stage6})
and exposes \code{POST /act} (raw state $\to$ obs $\to$ inference $\to$ action)
and \code{POST /reset} (clear hidden state, choose the shot target by strategy).
\textbf{Key decision:} the server computes its own observation (160-D
\code{compute_shooter_obs}) from raw MuJoCo state, so the observation space is
fully private to us -- \code{compete.py} passes only raw state and each team
computes its own observation, enabling interoperability across heterogeneous
systems.

\paragraph{Observation computation (\code{compute_shooter_obs}).} From the raw
JSON state, extract the shooter's 29 joint positions/velocities and last action,
base orientation and angular velocity, and the ball world position; transform the
ball and destination positions into the pelvis frame; read the reference joint
positions/velocities and torso angular velocity from the motion file at the
corresponding timestep; concatenate the nine terms into the 160-D observation.

\subsection{Three shooting strategies}
\begin{table}[h]
\centering\small
\caption{Shooting strategies.}
\begin{tabular}{@{}llp{6.4cm}l@{}}
\toprule
Strategy & Trigger & Behavior & Latency \\
\midrule
\code{uniform} & \code{/reset} & random motion (incl. left leg) $+$ random target $y\in[-1.4,1.4]$, locked immediately & none \\
\code{best-motion} & \code{/reset} & random target $+$ \code{_select_motion_for_dest()} picks the best right-leg motion, locked immediately & none \\
\code{gk-aware} & per-frame \code{/act} & observe the GK for 5--16 frames, then lock the far-side target $+$ best motion & 5--16 frames \\
\bottomrule
\end{tabular}
\end{table}

\subsection{The \code{gk-aware} strategy}
This is the default and most capable strategy, in three phases.

\textbf{Observation phase} (\code{shoot_planned["done"]==False}): each \code{/act}
appends the GK root-$y$ to \code{gk_y_history}; \code{_should_lock()} applies a
dual condition -- \emph{behavioral trigger} (GK lateral offset $\geq0.3$\,m for a
run of \code{min_observe}$=5$ frames $\to$ early lock) and \emph{hard deadline}
(reaching \code{deadline_pct}$=60\%$ of the kick frame $\to$ force lock).
Crucially we do \textbf{not} reset the policy on lock: locking only changes the
\code{dest_body} observation term and leaves the LSTM hidden state intact, because
the policy was trained on a continuous motion stream and jumping to frame 16 with
an empty hidden state causes an out-of-distribution collapse.

\textbf{Target planning} (\code{_plan_destination}): if the GK is left
($y<-0.3$), shoot right \code{random(0.3,1.4)}; if right ($y>+0.3$), shoot left
\code{random(-1.4,-0.3)}; if central ($|y|\leq0.3$), shoot the full goal width
\code{random(-1.4,1.4)}. The random component within the chosen half adds
unpredictability.

\textbf{Execution:} \code{_select_motion_for_dest(dest_y, motions)} selects the
best right-leg motion; the target world coordinate is set to $(-0.5,\text{dest}_y,
0.11)$ (goal-plane center); the motion continues to play and the policy steers the
kick from the target term.

\subsection{Target region $\to$ motion map}
Built from \code{eval_motion_analysis.py} (5000 trials per motion$\times$target):

\begin{table}[h]
\centering\small
\caption{Target region to best motion (data-driven).}
\begin{tabular}{@{}lllll@{}}
\toprule
Target $y$ range & Best motion & Speed & Accuracy & Reason \\
\midrule
$[-1.5,-0.7)$ far left & \code{006_right} & 16.2\,m/s & \textbf{0.10\,m} & best far-corner accuracy \\
$[-0.7,-0.3)$ mid-left & \code{009_right} & 14.2\,m/s & 0.35\,m & best right-leg mid-left accuracy \\
$[-0.3,+0.3]$ center & \code{010_right} & \textbf{16.4\,m/s} & 0.11\,m & highest speed $+$ good center accuracy \\
$(+0.3,+0.7]$ mid-right & \code{005_right} & 12.1\,m/s & \textbf{0.006\,m} & best accuracy overall (sub-cm) \\
$(+0.7,+1.5]$ far right & \code{005_right} & 15.7\,m/s & 0.64\,m & most versatile motion \\
\bottomrule
\end{tabular}
\end{table}
All selected motions are right-leg kicks. Left-leg motions are unused because
they are slower (6--9\,m/s) and completely fail the far-left corner (0\% goal
rate).

\subsection{Kick-frame map and zero-GK baseline}
\code{analyze_kick_timing.py} confirms all motions contact the ball at frames
26--30 (0.52--0.60\,s) with zero std, which makes the \code{_should_lock()}
deadline reliable (e.g.\ \code{001_right} kicks at frame 27/260,
\code{005_right} at 26/230, \code{006_right} at 27/225). When \code{compete.py}
omits \code{--goalkeeper-api}, the GK is a \code{ZeroPolicy} (holds the
\code{HOME_KEYFRAME} crouch at $y=0$), giving a pure technical baseline.

% ---------------------------------------------------------------------------
\section{Motion performance analysis}\label{sec:sh-motion}
% ---------------------------------------------------------------------------

\textbf{Scale:} 10 motions $\times$ 5 targets $\times$ 5000 trials $=
\mathbf{250{,}000}$ trials, model Stage 6 (\code{model_20000.pt}), targets
$y=-1.50,-0.75,0.00,+0.75,+1.50$ (even split of the 3.0\,m goal).

\paragraph{Overall.} Mean goal rate $\mathbf{94\%}$ (9/10 motions at 100\%); kick
trigger rate $100\%$; right-leg mean ball speed $14.2$\,m/s ($\approx51$\,km/h);
left-leg $8.0$\,m/s ($\approx29$\,km/h); right-leg mean accuracy $0.42$\,m;
left-leg $0.26$\,m.

\begin{table}[h]
\centering\small
\caption{Right-leg motion performance (all 100\% goal rate).}
\begin{tabular}{@{}lcccccc@{}}
\toprule
Motion & Frames & Mean speed & Mean accuracy & Hit@0.3m & Best target & Best accuracy \\
\midrule
\code{001_right} & 260 & 13.59\,m/s & 0.340\,m & 40\% & $+0.75$ & 0.152\,m \\
\code{004_right} & 257 & 14.02\,m/s & 0.548\,m & 20\% & $+0.75$ & 0.197\,m \\
\code{005_right} & 230 & 13.96\,m/s & 0.402\,m & 20\% & \textbf{$+0.75$} & \textbf{0.006\,m} \\
\code{006_right} & 225 & 14.56\,m/s & 0.343\,m & 40\% & \textbf{$-1.50$} & \textbf{0.095\,m} \\
\code{009_right} & 205 & 13.73\,m/s & 0.454\,m & 40\% & $+0.00$ & 0.144\,m \\
\code{010_right} & 230 & 15.40\,m/s & 0.512\,m & 20\% & $+0.00$ & 0.112\,m \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[h]
\centering\small
\caption{Left-leg motion performance (80\% goal rate; 0\% only at $y=-1.50$).}
\begin{tabular}{@{}lcccc@{}}
\toprule
Motion & Frames & Mean speed & Mean accuracy & Hit@0.3m \\
\midrule
\code{002_left} & 214 & 7.27\,m/s & 0.245\,m & 80\% \\
\code{003_left} & 289 & 7.88\,m/s & 0.277\,m & 65\% \\
\code{007_left} & 229 & 8.42\,m/s & 0.266\,m & 60\% \\
\code{008_left} & 225 & 8.36\,m/s & 0.261\,m & 60\% \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[h]
\centering\small
\caption{Target $\times$ ball-speed matrix (m/s), right-leg motions.}
\begin{tabular}{@{}lccccc@{}}
\toprule
Motion & $y{=}{-}1.50$ & $y{=}{-}0.75$ & $y{=}{+}0.00$ & $y{=}{+}0.75$ & $y{=}{+}1.50$ \\
\midrule
\code{001_right} & 14.58 & 14.04 & 14.14 & 12.41 & 12.80 \\
\code{004_right} & 13.02 & 14.52 & 14.62 & 12.53 & 15.41 \\
\code{005_right} & 14.59 & 13.37 & 14.12 & 12.06 & \textbf{15.69} \\
\code{006_right} & \textbf{16.18} & 14.18 & 14.04 & 12.99 & 15.42 \\
\code{009_right} & 14.40 & 14.24 & 12.14 & 12.54 & 15.33 \\
\code{010_right} & 14.64 & 14.57 & \textbf{16.42} & 16.28 & 15.10 \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Key findings.} (1) \emph{Strong right/left asymmetry:} right-leg ball
speed is $\approx2\times$ the left, and the left leg cannot reach the far-left
corner -- likely because right-leg motions are more numerous/better in the data
or the reward structure favors the right leg. (2) \emph{Far-corner accuracy
degrades:} all motions lose accuracy at $y=\pm1.50$ (0.4--1.2\,m vs.\ 0.1--0.4\,m
at center), a reasonable mechanical limit. (3) \emph{Competition recommendation:}
use right-leg motions only, with the \code{gk-aware} strategy that reads the GK
during execution and shoots to the far side.

% ---------------------------------------------------------------------------
\section{Engineering decisions and lessons}\label{sec:sh-engineering}
% ---------------------------------------------------------------------------

\paragraph{Unified architecture.} All six stages share an identical LSTM
(128-64-32 $+$ 2$\times$128), differing only by environment config. This lets
checkpoints transfer seamlessly across stages (\code{--resume} loads
actor$+$critic$+$normalizer), concentrates effort on reward/environment design
rather than re-tuning networks, and speeds iteration (changing a reward weight
needs no network redesign).

\paragraph{Coordinate alignment is a fatal detail.} Stage 5 switching the
training frame fully to \code{compete.py}'s world coordinates
(robot $(4,0,0.8)$, yaw $\pi$) looks trivial but is decisive: it removes the
sim-to-sim domain gap. Checkpoints trained earlier in the training frame degrade
markedly in the compete scene.

\paragraph{LSTM state continuity.} Not calling \code{policy.reset()} on target
lock in \code{gk-aware} was a key lesson. An initial implementation reset on
\code{should_lock}, feeding the LSTM an empty hidden state at motion frame 16 -- a
combination never seen in training -- causing an OOD collapse. The fix updates
only the \code{dest_body} term and preserves LSTM continuity.

\paragraph{Data-driven motion selection.} The \code{_select_motion_for_dest()}
map is not heuristic but grounded in the 250K-trial measurements; each target
region's recommendation is backed by concrete speed/accuracy numbers, avoiding
subjective guessing.

\paragraph{Rule-based vs.\ adversarial training.} Phase 2 pits the shooter against
an unknown opponent GK. Rather than \emph{adversarial training} (which is
unstable when two policies learn online and would compound an already complex
six-stage curriculum), we chose a \emph{strong single agent $+$ rule-based target
selection}: first train a near-perfect GK-free shooter (99.96\% goal rate), then
add a \code{gk-aware} rule that reads the GK and shoots to the far side. ``Shoot
to the far side'' is a physically deterministic decision that needs no learning,
and given a shooter that scores 12--16\,m/s anywhere in the goal, this beats a
random target by a wide margin against a reactive GK. The trade-off is reduced
adaptivity to complex GK behavior (occlusion, tackles), but interpretability and
controllability are preferable against an unknown opponent to a possibly
under-trained adversarial policy.

% ===========================================================================
\begin{thebibliography}{9}
% ===========================================================================
\bibitem{kong} J.~Kong, X.~Liu, Y.~Lin, J.~Han, S.~Schwertfeger, C.~Bai, and
X.~Li, ``Learning soccer skills for humanoid robots: A progressive
perception-action framework,'' 2026. arXiv:2602.05310.

\bibitem{ren} J.~Ren, J.~Long, T.~Huang, H.~Wang, Z.~Wang, F.~Jia, W.~Zhang,
J.~Wang, P.~Luo, and J.~Pang, ``Humanoid goalkeeper: Learning from
position-conditioned task-motion constraints,'' arXiv:2510.18002, 2025.

\bibitem{ppo} J.~Schulman, F.~Wolski, P.~Dhariwal, A.~Radford, and O.~Klimov,
``Proximal policy optimization algorithms,'' arXiv:1707.06347, 2017.

\bibitem{dagger} S.~Ross, G.~J.~Gordon, and J.~A.~Bagnell, ``A reduction of
imitation learning and structured prediction to no-regret online learning,'' in
\emph{Proc. AISTATS}, 2011, pp.~627--635.

\bibitem{icem} C.~Pinneri, S.~Sawant, S.~Blaes, J.~Achterhold, J.~Stueckler,
M.~Rolinek, and G.~Martius, ``Sample-efficient cross-entropy method for real-time
planning,'' in \emph{Proc. CoRL}, PMLR vol.~155, 2021, pp.~1049--1065.

\bibitem{cem} R.~Y.~Rubinstein and D.~P.~Kroese, \emph{The Cross-Entropy Method}.
New York: Springer, 2004.

\bibitem{mujoco} E.~Todorov, T.~Erez, and Y.~Tassa, ``MuJoCo: A physics engine
for model-based control,'' in \emph{Proc. IEEE/RSJ IROS}, 2012, pp.~5026--5033.

\bibitem{rudin} N.~Rudin, D.~Hoeller, P.~Reist, and M.~Hutter, ``Learning to walk
in minutes using massively parallel deep reinforcement learning,'' in
\emph{Proc. CoRL}, PMLR vol.~164, 2022, pp.~91--100.
\end{thebibliography}

\end{document}
